Draft Documents

### 2022-11-14 src/draft.Rmd

needs("dplyr", "MASS", "knitr", "readr", "tidyr", "ggplot2", "e1071", "moments", "corrplot", "Hmisc", "PerformanceAnalytics", "mice")

@@

corWithPov <- function(indevar, data) {
  cor(data[[indevar]], data[["pov"]], use = "complete.obs")
}

@@

@@

### 3.2.4. Imputation (To be done)
What is Imputation?
**1. What is imputation**  
Imputation is a technique to handle missing values by replacing missing data with substitute values. In our study, we focus on "item imputation", which means substituting for a component of a data point (i.e. a variable). The general idea is to take a value that preserve the property of the data (e.g., distribution, mean, standard deviation), or predict the value using other variables.

@@

```{r}
set.seed(1984)
# split data
idx <- sample(1:nrow(countries1), replace = F, 0.7 * nrow(countries1))
countries.train.4 <- countries1[idx,]
countries.test.4 <- countries[-idx,]

@@

# run the algorithm
# countries1.imputed <- mice(countries1, m = 5, maxit = 30, method = meth)
countries1.imputed <- readRDS("../countries1.imputed.RData")
summary(countries1.imputed)
```
We can take a look into one of the imputated data sets.
```{r}
countries1.imputed.1 <- complete(countries1.imputed, 1)
summary(countries1.imputed.1)
sum(!complete.cases(countries1.imputed.1))
```
We found no missing cases as expected.

#### 3.2.4.1. Model Fitting
We generated 5 sets of imputated (training) data. We can build our model on each of these data sets and combine the estimates using [pooling rule](https://amices.org/mice/reference/pool.html#details). We remove `reg` and `country.name` from the model because they're perfectly predicted by `country.code`

We can start with using one of the imputated sets.
```{r}
fit1 <- lm(pov ~ .-country.name, countries1.imputed.1)
summary(fit1)
```

We might be more interested with large regions of countries rather than individual countries.
```{r}
fit1reg <- lm(pov ~ .-country.name-country.code, countries1.imputed.1)
summary(fit1reg)
```

With all the generated data sets.

```{r}
fit2 <- with(data = countries1.imputed, exp = lm(pov ~ reg+year+income+edu.total+hlth+mil+fdi+lbr.part+unemp+pop.gwth.total+pop.gwth.rural+pop.gwth.urban+gdp.dflt+gcf+trade+gdp.pc+lfdi+lgdp.dflt+lgdp.pc))
fit2.combined <- pool(fit2)
summary(fit2.combined)
pool.r.squared(fit2.combined)
```

We can conduct a step-wise AIC variable selection. It is similar to the procedure we use in class, but based on a metric call AIC (Akaike Information Criterion), which is an estimator of prediction error and relative quality of statistical models.

continue with fitting and displaying model [link](https://thomaselove.github.io/432-notes/multiple-imputation-and-linear-regression.html#fit-linear-regression-using-ols-and-fit.mult.impute)  
-> AIC on one model [link](https://rforpoliticalscience.com/2020/10/23/choose-model-variables-by-aic-in-a-stepwise-algorithm-with-the-mass-package-in-r/)  
-> stepwise selection on mice  [link](https://stefvanbuuren.name/fimd/sec-stepwise.html)

>>>>

needs("dplyr", "MASS", "knitr", "readr", "tidyr", "ggplot2", "e1071", "moments", "corrplot", "Hmisc", "PerformanceAnalytics", "mice", "car", "glmnet")

@@

# correlation of independent variable and pov
corWithPov <- function(indevar, data) {
  cor(data[[indevar]], data[["pov"]], use = "complete.obs")
}

@@

```{r}
countries1 <- countries1 %>% 
  mutate(lfdi = log(fdi), lgdp.dflt = log(gdp.dflt), lgdp.pc = log(gdp.pc)) %>% 
  mutate(lfdi = ifelse(is.nan(lfdi) | lfdi == -Inf, NA, lfdi))
```

@@

### 3.2.4. Imputation
**1. What is imputation**  
Imputation is a technique to handle missing values by replacing missing data with substitute values. In our study, we focus on "item imputation", which means substituting for a component of a data point (i.e. a variable). The general idea is to take a value that preserve the property of the data (e.g., distribution, mean, standard deviation), or predict the value using other variables.

@@

```{r}
set.seed(1984)
# split data
isComplete <- which(complete.cases(countries1))
idx <- sample(isComplete, replace = F, 0.3 * nrow(countries1))
countries.train.4 <- countries1[-idx,]
countries.test.4 <- countries1[idx,]

@@

# run the algorithm
# countries1.imputed <- mice(countries1, m = 3, maxit = 20, method = meth)
# saveRDS(countries1.imputed, 'countries1.imputed.RData')
countries1.imputed <- readRDS('countries1.imputed.RData')
summary(countries1.imputed)
```
We can take a look into one of the imputated data sets.
```{r}
countries1.imputed.1 <- complete(countries1.imputed, 1)
summary(countries1.imputed.1)
sum(!complete.cases(countries1.imputed.1))
```
We found no missing cases as expected.

#### 3.2.4.1. Model Fitting
##### A. Ordinary Linear Regression
We generated 5 sets of imputated (training) data. We can build our model on each of these data sets and combine the estimates using [pooling rule](https://amices.org/mice/reference/pool.html#details). With all the generated data sets.

```{r}
formula.str <- 'pov ~ reg+year+income+edu.total+hlth+mil+fdi+lbr.part+unemp+pop.gwth.total+pop.gwth.rural+pop.gwth.urban+gdp.dflt+gcf+trade+gdp.pc+lfdi+lgdp.dflt+lgdp.pc'

# Using only the first imputed data set
fit1 <- lm(as.formula(formula.str), data = complete(countries1.imputed, 1))
summary(fit1)
fit1.pred <- predict(fit1, countries.test.4)

# Build models with all generated data set and pool the estimates
fit2 <- with(data = countries1.imputed, exp = lm(as.formula(formula.str)))
fit2.combined <- pool(fit2)
summary(fit2.combined)
# dummy lm model
fit2.dummy <- lm(as.formula(formula.str), data = countries1)
# replace coefficients of dummy model & predict
fit2.dummy$coefficients <- fit2.combined$pooled$estimate
fit2.pred <- predict(fit2.dummy, countries.test.4)
```
Helper function to calculate R-Squared
```{r}
r2 <- function(pred, orig) {
   RSS <- sum((pred - orig)^2)
   TSS <- sum((pred - mean(orig))^2)
   R2 <- 1 - RSS/TSS
   return (R2)
}

adjR2 <- function(pred, orig, k) {
  R2 <- r2(pred, orig)
  n <- length(pred)
  adjr2 <- 1 - (1-R2)*(n-1)/(n-k-1)
  return (adjr2)
}
```
Adjusted R-Squared on train and test
```{r}
k <- length(fit1$coefficients) - 1
fit1.train.r2 <- summary(fit1)$adj.r.squared
fit1.test.r2 <- adjR2(fit1.pred, countries.test.4$pov, k)
fit2.train.r2 <- pool.r.squared(fit2, adjusted = T)[,'est']
fit2.test.r2 <- adjR2(fit2.pred, countries.test.4$pov, k)
res <- data.frame(model = c("Single", "Pooled"), train = c(fit1.train.r2, fit2.train.r2), test = c(fit1.test.r2, fit2.test.r2))
kable(res)
```

The models seem to be over-fitting. We should perform model simplification.

##### B. Regularization

Prepare data

```{r}
sd0 <- function(vct) {
  if (!is.numeric(vct)) {
    return(NA)
  }

  return(sd(vct, na.rm = T))
}

std0 <- function(vct, scl) {
  if (!is.numeric(vct)) {
    return(vct)
  }

  return(vct/scl)
}

standardise <- function(train, test) {
  data <- rbind(train, test)
  scaler <- unlist(lapply(data, sd0))

  numCols <- which(unlist(lapply(data, is.numeric)))
  num <- as.data.frame(mapply(std0, vct = data[,numCols], scl = scaler[numCols], SIMPLIFY = T))
  fct <- data[,-numCols]
  return(cbind(fct, num))
}
```

```{r}
optimalModel <- function(train) {
  #standardized
}
```

##### B. Step-wise AIC
We can conduct a step-wise AIC variable selection. It is similar to the procedure we use in class, but based on a metric call AIC (Akaike Information Criterion), which is an estimator of prediction error and relative quality of statistical models. The lower AIC is, the better the model fits. \\
We have no idea how to pool coefficients for AIC model,

```{r}
# helper
fitAIC <- function(i) {
  fit <- lm(as.formula(formula.str), complete(countries1.imputed, i))
  aic.fit <- stepAIC(fit, trace = F, direction = 'backward')
  return(aic.fit)
}

# build models
aic.fits <- lapply(1:countries1.imputed$m, fitAIC)
# predict with each model
aic.pred <- lapply(aic.fits, predict, newdata = countries.test.4)
# calculate adj r2
aic.train.r2 <- unlist(lapply(aic.fits, \(model) summary(model)$adj.r.squared))

k <- lapply(aic.fits, \(m) length(m$coefficients))
aic.test.r2 <- unlist(mapply(adjR2, aic.pred, k = k, MoreArgs = list(orig = countries.test.4$pov)))

res <- data.frame(`set no.` = 1:3, train = aic.train.r2, test = aic.test.r2)
```

Adjusted R-Squared

```{r}
kable(res)
```

-> AIC on one model [link](https://rforpoliticalscience.com/2020/10/23/choose-model-variables-by-aic-in-a-stepwise-algorithm-with-the-mass-package-in-r/)

-> stepwise selection on mice  [link](https://stefvanbuuren.name/fimd/sec-stepwise.html)
